Supplementary materials: $R^*$ convergence diagnostic

### 2020-10-22 write_up/arxiv_submission_aug_2020/supplementary_materials_marked.tex
added trends

\subsection{250-dimensional model}
In the first of these, we used the 250-dimensional multivariate normal of eq. \eqref{eq:mvt_normal_250} with 250 post-warm-up iterations (after 250 warm-up iterations) for each of 4 chains from Stan's NUTS to calculate $R^*$ distributions as in Algorithm \ref{alg:R_star_uncertainty}. Here, we considered both the centered and non-centered parameterisations, where, in both cases, the number of iterations is comparable to the number of parameters, so the training data is relatively ``wide'': the $R^*$ distribution in each case is shown in Fig. \ref{fig:mvt_wide_both}A. In this figure, it is clear that both $R^*$ distributions are shifted away from 1, indicating non-convergence. The same conclusion is reached if rank-normalised split-$\widehat{R}$ is used instead (Fig. \ref{fig:wide_both_diagnostics}A), since, for both parameterisations, some of the parameters had $\widehat{R}>1.01$. Using bulk- or tail-ESS instead, we conclude that the non-centered parameterisation shows signs of convergence whereas the centered does not (Fig. \ref{fig:wide_both_diagnostics}B\&C).

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{mvt_wide_both.pdf}}
	\caption{\textbf{Wide data examples.} A shows the $R^*$ distribution for the 250-dimensional example in \S\ref{sec:wide} with 250 post-warm-up iterations per chain from Stan's NUTS algorithm across both model parameterisations; B shows $R^*$ distribution for the 10,000-dimensional example with 400 and 1000 MCMC iterations per chain (although the first half of these were discarded as warm-up). In all cases, 1000 draws of $R^*$ are plotted as generated by Algorithm \ref{alg:R_star_uncertainty} for a single MCMC run composed of 4 chains.}
	\label{fig:mvt_wide_both}
\end{figure}

@@

\subsection{10,000-dimensional model}
We next consider a more challenging example -- a target distribution with 10,000 dimensions. In this case, we assume independent standard normals for each dimension. In Fig. \ref{fig:mvt_wide_both}B, we plot the $R^*$ distribution for two MCMC runs targeting this distribution: one with 400 iterations, the other with 1000. In both cases, the distributions were right of $R^*=1$, indicating non-convergence. These results were also echoed by rank-normalised split-$\widehat{R}$, with 65\% of dimensions having $\widehat{R}>1.01$ for the 400 iteration case and 19\% for the 1000 iteration case.

>>>>

\subsection{250-dimensional model}
In the first of these, we used the 250-dimensional multivariate normal of eq. \eqref{eq:mvt_normal_250} with 250 post-warm-up iterations (after 250 warm-up iterations) for each of 4 chains from Stan's NUTS to calculate $R^*$ distributions as in Algorithm \ref{alg:R_star_uncertainty}. Here, we considered both the centered and non-centered parameterisations, where, in both cases, the number of iterations is comparable to the number of parameters, so the training data is relatively ``wide''. \textcolor{red}{We also calculated $R^*$ distributions using both the GBM and RF classifiers: the $R^*$ distribution in each case is shown in Figs. \ref{fig:mvt_wide_both}A\&B. This figure shows that, across both parameterisations, convergence has not yet been reached since all $R^*$ distributions are shifted rightwards of 1.} The same conclusion is reached if rank-normalised split-$\widehat{R}$ is used instead (Fig. \ref{fig:wide_both_diagnostics}A), since, for both parameterisations, some of the parameters had $\widehat{R}>1.01$. Using bulk- or tail-ESS instead, we conclude that the non-centered parameterisation shows signs of convergence whereas the centered does not (Fig. \ref{fig:wide_both_diagnostics}B\&C).

\begin{figure}[!htb]
	\centerline{\includegraphics[width=1\textwidth]{mvt_wide_both.pdf}}
	\caption{\textbf{Wide data examples.} Top row shows the $R^*$ distribution for the 250-dimensional example in \S\ref{sec:wide} with 250 post-warm-up iterations per chain from Stan's NUTS algorithm across both model parameterisations; bottom row shows $R^*$ distribution for the 10,000-dimensional example with 400 and 1000 MCMC iterations per chain (although the first half of these were discarded as warm-up). In the left column, results for the GBM classifier are shown; in the right, we show the same for the RF classifier. In all cases, 1000 draws of $R^*$ are plotted as generated by Algorithm \ref{alg:R_star_uncertainty} for a single MCMC run composed of 4 chains.}
	\label{fig:mvt_wide_both}
\end{figure}

@@

\subsection{10,000-dimensional model}
We next consider a more challenging example -- a target distribution with 10,000 dimensions. In this case, we assume independent standard normals for each dimension. \textcolor{red}{In Figs. \ref{fig:mvt_wide_both}C\&D, we plot the $R^*$ distributions from GBM and RF classifiers for two MCMC runs targeting this distribution: one with 400 iterations, the other with 1000.} In all cases, the distributions were right of $R^*=1$, indicating non-convergence. These results were also echoed by rank-normalised split-$\widehat{R}$, with 65\% of dimensions having $\widehat{R}>1.01$ for the 400 iteration case and 19\% for the 1000 iteration case.
